RickView: Lightweight Standalone Knowledge Graph Browsing

### 2023-12-05 paper.tex
more paper

\documentclass[runningheads]{llncs}
%\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{csquotes}
\usepackage{siunitx}
\usepackage{hyperref}
\renewcommand\UrlFont{\color{blue}\rmfamily}
\newcommand\citep[1]{\cite{#1}}
\urlstyle{rm}
\hypersetup{pdfauthor={Konrad Höffner},pdftitle={RickView: Lightweight Standalone Knowledge Graph Browsing Powered by Rust}}

\begin{document}
\title{RickView: Lightweight Standalone Knowledge Graph Browsing Powered by Rust}
%\titlerunning{Abbreviated paper title}
\author{Konrad Höffner\inst{1}\orcidID{0000-0001-7358-3217}}
\authorrunning{K. Höffner}
\institute{Institute for Medical Informatics, Statistics and Epidemiology (IMISE), Leipzig University, Leipzig, Germany
\email{konrad.hoeffner@uni-leipzig.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The standard backends for RDF browsers are SPARQL endpoints, which allow expressive queries, but they are overengineered for the simple task of browsing.
RickView is an RDF browser that follows an alternative approach of directly querying an in-memory dataset bypassing SPARQL.
It is implemented in Rust, which enables a high level of performance comparable to C and C++ while being memory-safe and thread-safe by design.
RickView is published as a single binary, optionally wrapped in a compact container image, without dependencies on external components.
It is adaptable to a given knowledge base via defaults that can be changed in a configuration file and overridden with environment variables.
It supports the binary RDF format Header Dictionary Triples (HDT), which allows querying compressed RDF in memory.
These features enable the stable long-term, low-maintenance, robust, simple and fast publication of multiple RDF knowledge graphs on low-end hardware.
\keywords{Linked Data browser \and RDF visualization.}
\end{abstract}

@@

\begin{figure}
\includegraphics{img/relatedwork.pdf}
\caption{Source: \cite{adaptinglodview}.
\label{fig:relatedwork}
\end{figure}

Search, browsing and exploration is a critical area in the linked data (LD) life cycle.
Similar to web browsers, where users enter the URL of a website and receive a display of an HTML document, in RDF browsers,
users enter the URL of an RDF resource and receive a human readable representation.
However in contrast to websites, RDF resources are intended to be processed by machines and do not provide a standardized way of being displayed.
\citep{adaptinglodview}
%An RDF graph, also called a knowledge graph, property graph or knowledge base, is a set of RDF triples in the form of \emph{(subject, predicate, object)} and typically describes a common domain under a shared namespace, which can be abbreviated with a prefix, such as \texttt{rdf}.
%For example, the RDF vocabulary has a \enquote{hash namespace} (ending with a \texttt{\#}) of \texttt{http://www.w3.org/1999/02/22-rdf-syntax-ns\#},
%which signifies that all resources, like \texttt{rdf:type} and \texttt{rdf:property}, of the vocabulary are described in the same document.
%This is suitable for small graphs that are not intended to be viewed by non-expert users.

@@

>>>>

\documentclass[runningheads]{llncs}
%\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{csquotes}
\usepackage{siunitx}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{booktabs}
\renewcommand\UrlFont{\color{blue}\rmfamily}
\newcommand\citep[1]{\cite{#1}}
\urlstyle{rm}
\hypersetup{pdfauthor={Konrad Höffner},pdftitle={RickView: Lightweight Standalone Knowledge Graph Browsing Powered by Rust}}

\begin{document}
\title{RickView: Lightweight Standalone Knowledge Graph Browsing}
%\titlerunning{Abbreviated paper title}
\author{Konrad Höffner\inst{1}\orcidID{0000-0001-7358-3217}}
\authorrunning{K. Höffner}
\institute{Institute for Medical Informatics, Statistics and Epidemiology (IMISE), Leipzig University, Leipzig, Germany
\email{konrad.hoeffner@uni-leipzig.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The standard backends for RDF browsers are SPARQL endpoints, which allow expressive queries, but they are overengineered for the simple task of browsing.
RickView is an URI dereferenciation tool and RDF browser that follows an alternative approach of directly querying an in-memory dataset bypassing SPARQL.
It is implemented in Rust, which enables a high level of performance comparable to C and C++ while being memory-safe and thread-safe by design.
RickView is published as a single binary, optionally wrapped in a compact container image, without dependencies on external components.
It is adaptable to a given knowledge base via defaults that can be changed in a configuration file and overridden with environment variables.
It supports the binary RDF format Header Dictionary Triples (HDT), which allows querying compressed RDF in memory.
These features enable the stable long-term, low-maintenance, robust, simple and fast publication of multiple RDF knowledge graphs on low-end hardware.
\keywords{Linked Data browser \and RDF visualization.}
\end{abstract}

@@

Search, browsing and exploration is a critical area in the linked data (LD) life cycle.
Similar to web browsers, where users enter the URL of a website and receive a display of an HTML document, in RDF browsers,
users enter the URL of an RDF resource and receive a human readable representation.
However in contrast to websites, RDF resources are intended to be processed by machines and do not provide a standardized way of being displayed.
%An RDF graph, also called a knowledge graph, property graph or knowledge base, is a set of RDF triples in the form of \emph{(subject, predicate, object)} and typically describes a common domain under a shared namespace, which can be abbreviated with a prefix, such as \texttt{rdf}.
%For example, the RDF vocabulary has a \enquote{hash namespace} (ending with a \texttt{\#}) of \texttt{http://www.w3.org/1999/02/22-rdf-syntax-ns\#},
%which signifies that all resources, like \texttt{rdf:type} and \texttt{rdf:property}, of the vocabulary are described in the same document.
%This is suitable for small graphs that are not intended to be viewed by non-expert users.

@@

\section{Related Work}\label{relatedwork}
Among the various kinds of LOD visualization tools (see \cref{fig:relatedwork}), RickView is an \emph{RDF and LOD browser} that fulfills three of the four criteria defined in~\cite{adaptinglodview} for a \emph{Pubby-class tool}:
It is (1) a URI dereferencing tool, (2) a LOD visualization tool and (3) an RDF browser.
However by design it is \emph{not} (4) a SPARQL endpoint interface because it does not retrieve information about resources from a SPARQL endpoint but instead from its in-memory representation of the dataset.
%Among the listed 
\citep{adaptinglodview} identifies limitations of LodView which are partially fixed in a unmaintained fork~\footnote{\url{https://github.com/ManlyMan1/LodView_Cyrillic}} with the last change being from 2020.
A pull request to integrate into the original LodView\footnote{\url{https://github.com/LodLive/LodView}} was originally planned but has not yet been created.

\begin{tabular}{lll}
\toprule
Feature												&LodView	&RickView\\
Resolution of Cyrillic URIs							&\\
Decoding Cyrillic URIs in Turtle representations	&\\
Support of Cyrillic literals						&\\
User-friendly URLs for RDF representations			&&\\
Support of hash URIs			&&\\
Expanding nested resources			&&\\
Support of RDF collections			&&\\
Pagination of resource property values			&&\\
Support of LATEX math notation			&&\\
\midrule
\bottomrule
\end{tabular}

%\begin{enumerate}
%\item a URI dereferenciation tool, i.e. when a client requests the URI of an RDF resource via HTTP protocol, the tool handles the request and responds with a resource representation;
%\item a LOD visualization tool, i.e. it represents the resource in human-friendly visual form;
%\item a Linked Data or RDF browser, i.e. it allows a user to navigate between RDF resources of an RDF dataset or the entire LOD cloud, by following links in resource representations;
%\end{enumerate}

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{img/relatedwork.pdf}
\caption{Categorization of LOD visualization tools. Source: \cite{adaptinglodview}.}
\label{fig:relatedwork}
\end{figure}
